---
title: "P8105_hw6_zo2168"
author: "Zhengkun Ou"
date: "2024-11-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(modelr)
library(mgcv)
library(SemiPar)
```

## Question 1
```{r}

```

## Question 2

```{r}
homicide <- read_csv(file = "homicide-data.csv", na = c(".", "NA", ""))
homicide_clean <- 
  homicide |> 
  janitor::clean_names() |>
  mutate(city_state = str_c(city, state, sep = ", "), 
         homicide_solved = ifelse(disposition %in% c("Closed without arrest", "Closed by arrest"), TRUE, FALSE),
         victim_age = ifelse(victim_age == "Unknown", NA, as.numeric(victim_age))) |>
  filter(!city_state %in% c("Dallas, TX","Phoenix, AZ","Kansas City, MO", "Tulsa, AL"), 
         victim_race %in% c("White", "Black"))

```

```{r}
baltimore_df = 
  homicide_clean |>
  filter(city_state == "Baltimore, MD")
  
baltimore_model <- glm(homicide_solved ~ victim_age + victim_sex + victim_race, data = baltimore_df)

baltimore_results = broom::tidy(baltimore_model, conf.int = TRUE) |>
                    filter(term == "victim_sexMale") |>
                    mutate(OR = exp(estimate), 
                           OR_lower = exp(conf.low),
                           OR_upper = exp(conf.high)) |> 
                    select(OR, OR_lower, OR_upper)

city_results = homicide_clean |> 
  nest(data = -city_state) |>
  mutate(
    models = map(data, \(x) glm(homicide_solved ~ victim_age + victim_sex + victim_race, 
                           data = x)),
    results = map(models, \(x) broom::tidy(x, conf.int = TRUE))
  ) |>
  unnest(results) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    OR_lower = exp(conf.low),
    OR_upper = exp(conf.high)
  ) |>
  select(city_state, OR, OR_lower, OR_upper)
  
  
city_results |>
 mutate(city_state = fct_reorder(city_state, OR)) |>
 ggplot(aes(x = city_state, y = OR)) +
 geom_point() + 
 geom_errorbar(aes(ymin = OR_lower, ymax = OR_upper)) + 
 coord_flip() +
 geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
 labs(
   title = "Odds Ratios for Solving Homicides (Male vs Female Victims)",
   x = "City",
   y = "Odds Ratio (Male vs Female)"
 )
```
Most cities have OR < 1 (to the left of red dashed line), meaning that homicides with male victims are generally less likely to be solved than those with female victims.

Cities range in ORs from about 0.6 (New York) to 1.2 (Fresno), showing considerable variation.

Some cities (like Philadelphia, Chicago) have narrow CIs, suggesting more precise estimates
Others (like Fresno, San Bernardino) have wide CIs, indicating less precision

Many cities have CIs that don't cross 1, suggesting significantly lower odds of solving male victim homicides
A few cities (like Fresno, Minneapolis) have CIs crossing 1, meaning the male/female difference isn't statistically significant

The ordering of cities by OR helps visualize the pattern and identify potential regional or size-based trends.


## Problem 3

```{r}
birth_weight = read_csv("birthweight.csv", na = c(".", "NA", "")) |> janitor::clean_names()
skimr::skim(birth_weight)
birth_weight_clean = birth_weight |> mutate(babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
                                            frace = factor(frace, levels = c(1, 2, 3, 4, 8), 
                                                          labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
                                            malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
                                            mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                                                          labels = c("White", "Black", "Asian", "Puerto Rican", "Other")))

```
```{r}
lm_1 <- lm(bwt ~ babysex + bhead + blength + gaweeks + 
                   delwt + smoken + wtgain + ppbmi, 
                   data = birth_weight_clean)
lm_1_tidy <- broom::tidy(lm_1)
summary(lm_1)
lm_1_tidy |> 
```
```{r}
birth_weight_clean |>
  add_predictions(lm_1) |>
  add_residuals(lm_1) |>
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, color = "blue")

```
```{r}
lm_2 <- lm(bwt ~ blength + gaweeks, data = birth_weight_clean)
lm_3 <- lm(bwt ~ bhead * blength * babysex, data = birth_weight_clean)

birth_df = 
  crossv_mc(birth_weight_clean, 100) |> 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
                                                  
birth_df = 
  birth_df |>
  mutate(model1 = map (train, \(df) lm(bwt ~ babysex + bhead + blength + gaweeks + 
                   delwt + smoken + wtgain + ppbmi, 
                   data = df)),
         model2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
         model3 = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))
  ) |>
  mutate(rmse_model1 = map2_dbl(model1, test, \(mod, df) rmse(model = mod, data = df)), 
         rmse_model2 = map2_dbl(model2, test, \(mod, df) rmse(model = mod, data = df)),
         rmse_model3 = map2_dbl(model3, test, \(mod, df) rmse(model = mod, data = df)))
  
birth_df |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse", 
    names_prefix = "rmse_"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() +
  labs(
    title = "Cross-validated RMSE by Model",
    x = "Model",
    y = "RMSE"
  )

```

